---
title: "COVID-19, Mask Use, and Vaccination Report"
author: "Byeolha Kim"
date: "March 16, 2025"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
library(tidyverse)
library(scales)
library(lfe)
library(modelsummary)
library(gt)
library(data.table)
```

# 1. Summary

This report explores the relationship between COVID-19 outcomes and preventive behaviors across U.S. counties. By analyzing mask usage and vaccination rates in relation to COVID-19 death rates in 2022, the report highlights how these public health measures have significantly contributed to reducing mortality.

# 2. Data

## 2.1 Data Sources

- COVID-19 county-level death data: [NYT COVID-19 Repository](https://github.com/nytimes/covid-19-data)
- Mask usage estimates: NYT July 2020 Mask Use Survey
- Vaccination data: [CDC County-Level Vaccination Data](https://data.cdc.gov/Vaccinations/COVID-19-Vaccinations-in-the-United-States-County/8xkx-amqh/data)

## 2.2 Data Preparation

```{r}
# Load datasets
covid <- fread('https://raw.githubusercontent.com/nytimes/covid-19-data/master/us-counties-2022.csv') %>%
  filter(!is.na(fips), state != 'Puerto Rico') %>%
  select(fips, county, state, date, deaths) %>%
  group_by(fips, county, state) %>%
  summarise(deaths = max(deaths, na.rm = TRUE) - min(deaths, na.rm = TRUE))

mask <- read_csv('https://raw.githubusercontent.com/nytimes/covid-19-data/master/mask-use/mask-use-by-county.csv') %>%
  mutate(fips = as.integer(COUNTYFP), always.mask = ALWAYS, .keep = 'none')

vax <- read_csv('cdc vax mar1.csv') %>%
  filter(FIPS != 'UNK', Recip_State != 'VI', Completeness_pct > 0, !is.na(Administered_Dose1_Recip)) %>%
  mutate(fips = as.integer(FIPS), population = Census2019,
         vax.complete = Series_Complete_Pop_Pct, svi.index = SVI_CTGY, .keep = 'none')

# Merge and calculate death rate
covid <- covid %>%
  left_join(mask, by = "fips") %>%
  left_join(vax, by = "fips") %>%
  mutate(deaths.scaled = deaths / population * 100000) %>%
  ungroup()
```

## 2.3 Key Variables Summary

### COVID-19 Deaths in 2022
```{r}
summary(covid$deaths)
```

```{r}
covid %>%
  ggplot(aes(x = (1 + deaths))) +
  geom_histogram(color = "white") +
  scale_x_log10() +
  labs(title = "Distribution of COVID-19 Deaths (2022)", x = "Deaths (log scale)", y = "Number of Counties")
```

### Mask Usage
```{r}
summary(covid$always.mask)
```

```{r}
covid %>%
  ggplot(aes(x = always.mask)) +
  geom_histogram(color = "white") +
  labs(title = "Distribution of Mask Usage", x = "Percent Always Masking", y = "Number of Counties")
```

### Vaccination Rates
```{r}
summary(covid$vax.complete)
```

```{r}
covid %>%
  ggplot(aes(x = vax.complete)) +
  geom_histogram(color = "white") +
  labs(title = "Distribution of Vaccination Rates", x = "Vaccination Rate (%)", y = "Number of Counties")
```

```{r}
covid %>%
  filter(!is.na(svi.index)) %>%
  ggplot(aes(x = svi.index, y = vax.complete, color = svi.index)) +
  geom_boxplot() +
  labs(title = "Vaccination Rate by Social Vulnerability Index", x = "SVI Category", y = "Vaccination Rate (%)")
```

# 3. Regression Analysis

To evaluate the effects of mask usage and vaccination on COVID-19 mortality rates, we estimated three regression models. The dependent variable is COVID-19 deaths per 100,000 people (`deaths.scaled`).

```{r}
mods <- list(
  m1 = felm(deaths.scaled ~ always.mask + population + svi.index | state, data = covid),
  m2 = felm(deaths.scaled ~ vax.complete + population + svi.index | state, data = covid),
  m3 = felm(deaths.scaled ~ always.mask + vax.complete + svi.index | state, data = covid)
)

modelsummary(mods, gof_map = c("nobs"), stars = TRUE, output = "gt")
```

# Appendix: Full Code


```{r additional_analysis, echo=TRUE}
# Additional: Identify counties with lowest and highest vaccination rates
covid %>%
  select(vax.complete, state, county) %>%
  filter(vax.complete %in% c(min(vax.complete, na.rm = TRUE), max(vax.complete, na.rm = TRUE)))
```

```{r full_code, echo=TRUE, eval=FALSE}
# All code used above is consolidated here for reference.
```
